Bandit Multiclass Classification

### 2018-08-08 main.tex
.

%\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% For citations
\usepackage{natbib}
\usepackage{amsthm}
% For algorithms
%\usepackage[algcompatible]{algpseudocode}
%\usepackage{algorithm}
%\usepackage{algorithmicx}
%\usepackage[noend]{algcompatible}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amsmath}           
\usepackage{mathtools}    
\usepackage{amsopn}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
%%%%%%%%%%%%%%%%%% toc 
\usepackage{scrwfile}
\TOCclone[\contentsname~(\appendixname)]{toc}{atoc}
\newcommand\StartAppendixEntries{}
\AfterTOCHead[toc]{%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=-10000\relax}%
}
\AfterTOCHead[atoc]{%
  \edef\maintocdepth{\the\value{tocdepth}}%
  \value{tocdepth}=-10000\relax%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=\maintocdepth\relax}%
}
\newcommand*\appendixwithtoc{%
  \cleardoublepage
  \appendix
  \addtocontents{toc}{\protect\StartAppendixEntries}
  \listofatoc
}
\usepackage{blindtext}
\usepackage{color}
%%%%%%%%%%%%%%%%%% toc

@@

\newcommand{\calA}{{\mathcal{A}}}
\newcommand{\calX}{{\mathcal{X}}}
\newcommand{\calS}{{\mathcal{S}}}
\newcommand{\calI}{{\mathcal{I}}}
\newcommand{\calJ}{{\mathcal{J}}}
\newcommand{\calK}{{\mathcal{K}}}
\newcommand{\calD}{{\mathcal{D}}}
\newcommand{\calE}{{\mathcal{E}}}
\newcommand{\calR}{{\mathcal{R}}}
\newcommand{\calT}{{\mathcal{T}}}
\newcommand{\calH}{\mathcal{H}}
\newcommand{\avgR}{\wh{\cal{R}}}
\newcommand{\ips}{\wh{r}}
\newcommand{\whpi}{\wh{\pi}}
\newcommand{\whE}{\wh{\E}}
\newcommand{\whV}{\wh{V}}
\newcommand{\Reg}{\text{\rm Reg}}
\newcommand{\whReg}{\wh{\text{\rm Reg}}}
\newcommand{\flg}{\text{\rm flag}}
\newcommand{\one}{\boldsymbol{1}}
\newcommand{\var}{\Delta}
\newcommand{\p}{\prime}
\newcommand{\nb}{\nabla}
\newcommand{\e}{\mathbf{e}}

@@

%#############################################################################
%    Introduction
%#############################################################################

\section{A Regret Lower Bound for A Certain Type of Algorithms}
In this section, we try to construct an error lower bound for a certain type of algorithms. This type of algorithms does not make update when it makes a wrong prediction. For simplicity, we only consider binary classification.  More formally, the algorithms we consider satisfy the following assumption.

@@

\textbf{Discussion}. We believe that $V$ can be in the order of $\mathcal{O}\left(\gamma^d\right)$ (some weaker thing like $\mathcal{O}\left(\gamma^{d/2}\right)$ is also acceptable). Basically, we need to figure out the following question: can we construct a sequence of points $\{x_1,x_2, \ldots\}$ , such that for every $t$, $x_t$'s distance with $\text{conichull}\{x_1,\ldots, x_{t-1}\}$ is larger than $\gamma$, but the volume difference $v(\text{conichull}\{x_1,\ldots, x_{t}\}) - v(\text{conichull}\{x_1,\ldots, x_{t-1}\})$ is upper bounded by $V$?

This lower bound does not rule out those algorithms that change its probability vector based on the $\textbf{count}$ of consecutive errors. This type of algorithm may still easy to be implemented (like QBC). Also it does not rule out the banditron algorithm.

@@

\section{Cone Algorithm}
\begin{algorithm}[H]
\caption{Banditron}
\textbf{definition}: $K\triangleq$ number of classes, $\gamma\triangleq$  margin\\
\textbf{Initialize}: $\calS_1=\cdots=\calS_K=\phi$ (empty set) \\
\For{$t=1, \ldots, T$}{
    Receive $x_t\in \mathbb{R}^d$. \\
    Define the cone $\calC_i = \left\{x\in\mathbb{R}^d: x=\sum_{j=1}^{|\calS_i|} \alpha_jy_j, \text{\ where\ } y_j\in \calS_i, \alpha_j\geq 0 \right\}$ \\
    (that is, $\calC_i$ is the conic hull of $\calS_i$) \\
    Check whether $x_t$ belongs to, or has distance smaller than $\gamma$, to one of $\calC_1, \ldots, \calC_K$. \\
    If so, classify $x_t$ to the corresponding class (say class $i$), and let $\calS_i\leftarrow \calS_i \cup \{x_t\}$. {\color{red} This prediction will be correct for sure by our margin assumption.}\\
    If not, let $\tilde{y}_t\sim \text{unif}([K])$ and predict $\tilde{y}_t$. If $\tilde{y}_t=y_t$, then $\calS_{y_t}\leftarrow \calS_{y_t} \cup \{x_t\}$. 
}
\end{algorithm}

\section{One-against-all Perceptron}
Same as Chicheng's writeup's Section 5: Fixed-Threshold Perceptron.

@@

\newpage
%\bibliography{adaptivity} 
%\bibliographystyle{plain}

>>>>

%\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% For citations
\usepackage{natbib}
\usepackage{amsthm}
% For algorithms
%\usepackage[algcompatible]{algpseudocode}
%\usepackage{algorithm}
%\usepackage{algorithmicx}
%\usepackage[noend]{algcompatible}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amsmath}           
\usepackage{mathtools}    
\usepackage{amsopn}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{soul}
%%%%%%%%%%%%%%%%%% toc 
\usepackage{scrwfile}
\TOCclone[\contentsname~(\appendixname)]{toc}{atoc}
\newcommand\StartAppendixEntries{}
\AfterTOCHead[toc]{%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=-10000\relax}%
}
\AfterTOCHead[atoc]{%
  \edef\maintocdepth{\the\value{tocdepth}}%
  \value{tocdepth}=-10000\relax%
  \renewcommand\StartAppendixEntries{\value{tocdepth}=\maintocdepth\relax}%
}
\newcommand*\appendixwithtoc{%
  \cleardoublepage
  \appendix
  \addtocontents{toc}{\protect\StartAppendixEntries}
  \listofatoc
}
\usepackage{blindtext}
\usepackage{color}
%%%%%%%%%%%%%%%%%% toc

@@

\newcommand{\calA}{{\mathcal{A}}}
\newcommand{\calX}{{\mathcal{X}}}
\newcommand{\calS}{{\mathcal{S}}}
\newcommand{\calI}{{\mathcal{I}}}
\newcommand{\calJ}{{\mathcal{J}}}
\newcommand{\calK}{{\mathcal{K}}}
\newcommand{\calD}{{\mathcal{D}}}
\newcommand{\calE}{{\mathcal{E}}}
\newcommand{\calR}{{\mathcal{R}}}
\newcommand{\calT}{{\mathcal{T}}}
\newcommand{\calH}{\mathcal{H}}
\newcommand{\avgR}{\wh{\cal{R}}}
\newcommand{\calW}{{\mathcal{W}}}
\newcommand{\ips}{\wh{r}}
\newcommand{\whpi}{\wh{\pi}}
\newcommand{\whE}{\wh{\E}}
\newcommand{\whV}{\wh{V}}
\newcommand{\Reg}{\text{\rm Reg}}
\newcommand{\whReg}{\wh{\text{\rm Reg}}}
\newcommand{\flg}{\text{\rm flag}}
\newcommand{\one}{\boldsymbol{1}}
\newcommand{\var}{\Delta}
\newcommand{\p}{\prime}
\newcommand{\nb}{\nabla}
\newcommand{\e}{\mathbf{e}}

@@

\section{Problem Statement}
A multiclass linear classifier is defined by a matrix $W\in \mathbb{R}^{K\times d}$, where $K$ is the number of classes ($K\geq 3$) and $d$ is the dimension of the feature vector. A set of samples $\{(x_t, y_t)\}_{t=1}^T$ where $x_t\in \mathbb{R}^d$ and $y_t\in[K]$ is called \textit{linearly separable} if there exists a $W^*$ such that $ y_t= \argmax_{y\in[K]} (W^*x_t)_y $ for all $t\in[T]$. (Note that $(W^*x_t)_y$ can also be written as $\e_y^\top W^* x_t$)

A set of samples $\{(x_t, y_t)\}_{t=1}^T$ is called \textit{linearly separable with a margin $\gamma$} if there exists a $W^*$ with $\|\e_i^\top W^*\|_2\leq 1$ for all $i$ and $(W^*x_t)_{y_t} \geq \max_{y\neq y_t} (W^*x_t)_{y} + \gamma$ for all $t$. For simplicity, we define the set where $W^*$ lies in as $\calW = \{W\in \mathbb{R}^{K\times d}: \|\e_i^\top W^*\|_2\leq 1\}$.

Consider the following \textit{online multiclass classification} problem: at time $t$, the environment first reveals $x_t\in \mathbb{R}^d$ (and we assume $\|x_t\|_2\leq 1$), then the learner predicts some label $\tilde{y}_t\in [K]$; finally the environment reveals the true label $y_t\in[K]$. It is known that if the samples are linearly separable, then the learner will only make constant mistakes. It is achievable with the following simple perceptron algorithm:

\begin{algorithm}[H]
\caption{Perceptron}
$W_1=\mathbf{0}$\\
\For{$t=1$ to $T$}{
   Predict $\tilde{y}_t=\argmax_y (W_tx_t)_y$ \\
   Update $W_{t+1}\leftarrow W_t + (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top$. 
}
\end{algorithm}
The analysis is simple. On one hand we have (note $\langle A,B \rangle:=\text{Tr}(A^\top B)$):
\begin{align*}
    \langle W_t, W^* \rangle &= \langle W_{t-1}, W^* \rangle + \langle (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top, W^* \rangle \\
    &= \langle W_{t-1}, W^* \rangle + (\e_{y_t}^\top W^* x_t-\e_{\tilde{y}_t}^\top W^* x_t) \\
    &\geq \langle W_{t-1}, W^* \rangle + \gamma\one[\tilde{y}_t\neq y_t]. 
\end{align*}
By induction, we get $\langle W_{T+1}, W^* \rangle \geq \gamma\sum_{t=1}^T \one[\tilde{y}_t\neq y_t]$.

On the other hand, 
\begin{align*}
    \langle W_{T+1}, W^* \rangle \leq \|W_{T+1}\|_F\|W^*\|_F \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \|x_t\|_2^2}\times 1 \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] } 
\end{align*}
Combining the above two inequality, we get $ \sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \leq \frac{4K}{\gamma^2}$.

The \textit{bandit multiclass classification} problem proceeds in a very similar way. The only difference is that the learner only knows whether she predicts \textbf{correctly or not}, rather than gets the true label (i.e., in each round, the feedback is only $\one[\tilde{y}_t=y_t]$, rather than $y_t$). Surprisingly, with this extremely limited feedback, the learner can still make only constant mistakes. It can be achieved with the following variant of \textit{halving algorithm}:

\begin{algorithm}[H]
\caption{Halving}
Discretize the space of $W$ (i.e., the $\calW$ as defined in the beginning) with balls of radius $\frac{1}{2\gamma}$. Let the set of discretization points be $\calS$. Then this is saying that for all $W\in \calW$, there is always a $W'\in \calS$ such that $\|W-W'\|_F\leq \frac{1}{2\gamma}$.\\
Let $\calS_1=\calS$ \\
\For{$t=1$ to $T$}{
    Let $\calS_t(y) = \{W\in \calS_t: (Wx_t)_y \geq (Wx_t)_i \ \forall i\in[K]\}$ (i.e., the set of $W$'s in $\calS_t$ that predict $y$ as the label of $x_t$) \\
    Let $\tilde{y}_t = \argmax_y |\calS_t(y)|$ (i.e., pick the majority vote)\\
    \If{$\tilde{y}_t\neq y_t$}{
        $\calS_{t+1}=\calS_t \backslash \calS_t(\tilde{y}_t)$ (i.e., eliminate the $W$'s that are inconsistent with the outcomes)
    }
    \Else{
        $\calS_{t+1}=\calS_t$
    }
}
\end{algorithm}
\textbf{Analysis}.  The majority in $\calS_t$ always has cardinality no less than $\frac{1}{K}|\calS_t|$. So every time the learner predicts incorrectly, the size of $|\calS_t|$ shrinks by an order of $(1-\frac{1}{K})$. By our margin assumption, there is a $W'\in \calS$ which incurs no error in all samples. Therefore, $|S_t|\geq 1$ always holds. This means the number of errors is bounded by the order of $\frac{\log(|\calS|)}{-\log(1-\frac{1}{K})}\leq \mathcal{O}\left( K^2d\log \frac{1}{\gamma} \right)$.

The main issue of this halving algorithm is that it is \textbf{inefficient} in the sense that $|\calS|$ is in the order of $\frac{1}{\gamma^{Kd}}$.

Hence, we want to answer the following question: \ul{for bandit multiclass classification, can we have an efficient algorithm (time complexity polynomial in $K,d,\frac{1}{\gamma}, T$) that guarantees constant mistake bound (polynomial in $K,d,\frac{1}{\gamma}$) in the $\gamma$-separable case?} (An unsolved open problem in \cite{kakade2008efficient})

\section{Naive Approaches}
\begin{itemize}
    \item \textbf{Exhausting the feature space}: discretize the feature space into $ \left(\frac{1}{\gamma}\right)^{\mathcal{O}(d)} $ blocks. The margin assumption guarantees that two points in a block would have the same label (so we can give every block a label). Each time a new point $x_t$ comes, see whether the learner already knows the correct label for that block. If yes, predict that label; if not, randomly choose a label that has not been chosen for that block. \\
    $\Rightarrow $ error bound: $K\left(\frac{1}{\gamma}\right)^{\mathcal{O}(d)} $. 
    \item \textbf{Exhausting the hypothesis space}: Discretize the hypothesis space into $\left(\frac{1}{\gamma}\right)^{Kd}$ discretization points (like in the halving algorithm). For each of them, make predictions based on it until it makes an mistake. \\
    $\Rightarrow $ error bound: $\left(\frac{1}{\gamma}\right)^{Kd}$. 
\end{itemize}

\section{Difficulties}
\label{section: difficulty}
Efficient classification algorithms often optimize a convex surrogate loss rather than a 0-1 loss. Examples include: 
\begin{itemize}
    \item Hinge loss: $\ell_t(W) = [1-(Wx_t)_{y_t}+\max_{y\neq y_t}(Wx_t)_{y}]_+$. 
    \item Logistic loss: $\ell_t(W)=-\ln\frac{\exp((Wx_t)_{y_t})}{\sum_{y}\exp((Wx_t)_y)}$
    \item A family of loss that interpolates hinge loss and squared hinge loss (see \cite{beygelzimer2017efficient}'s Eq.(3))
\end{itemize}
In full-information setting, one can directly use online convex optimization techniques to deal with the above loss functions (learning over the space of $W$).

Many works for bandit classification reuse this kind of convex optimization schemes, together with explicit exploration and inverse propensity weighting \cite{kakade2008efficient, hazan2011newtron, beygelzimer2017efficient, foster2018logistic}. However, because it is hard to estimate these losses when the true label $y_t$ is not known, some of these algorithms \cite{beygelzimer2017efficient, foster2018logistic} simply do not update when the learner makes a mistake ($\tilde{y}_t \neq y_t$). We give this kind of algorithm an error lower bound $\Omega\left(\left(\frac{1}{\gamma}\right)^{(d-1)/2}\right)$ in Section~\ref{section: passive lower bound}.

It indeed looks hard to design a convex loss for $W$'s when the learner makes a mistake: when $\tilde{y}_t
\neq y_t$, the set of $W$'s that we want to penalize (i.e., to assign larger loss) is $\{W: (Wx_t)_{\tilde{y}_t} \geq (Wx_t)_{y}, \forall y \}$, which is a convex cone in the space of $W$. It is impossible for a convex function to be large only in a convex subset (the case $K=2$, on the other hand, does not have this issue). Can we argue that if the learner is restricted to use convex losses in the space of $W$, she will have to suffer exponential (in $K$) errors?

\section{One-versus-all Kernel Perceptron}
\subsection{One-versus-all Perceptron}
Our assumption of linearly separable with a margin is 
\begin{align}
 (W^*x_t)_{y_t} \geq (W^*x_t)_{y} + \gamma, \ \forall y\neq y_t. \label{eqn:weaker}
\end{align}
As discussed in Section~\ref{section: difficulty}, it seems hard to make update when $\tilde{y}_t\neq y_t$ and achieve a constant and polynomial error bound.

For a moment, in this subsection we make the following stronger margin assumption (which we call \textit{one-versus-all separable} with a margin): 
\begin{align}
\begin{cases}
    (W^*x_t)_{y_t} \geq \gamma/2 \\
    (W^*x_t)_y \leq -\gamma/2, \forall y\neq y_t.
\end{cases}
\label{eqn:stronger}
\end{align}
Clearly, one-versus-all separability implies linearly separability, but not the other way around. With one-versus-all separability assumption, we can view the problem as $K$ parallel binary classification problem. The following algorithm achieves constant error bound: 
\begin{algorithm}[H]
\caption{One-versus-all Perceptron}
\label{alg:ova}
Initialize $w_t^{(1)}=\ldots=w_t^{(K)}=\mathbf{0}\in \mathbb{R}^d$\\
\For{$t=1$ to $T$}{
   \If{$\exists y$ such that $w_t^{(y)\top} x_t \geq 0$}{
       Assign $\tilde{y}_t$ to any $y$ with $w_t^{(y)\top}  x_t \geq 0$ \\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t\neq y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} - x_t$ } \label{line:update1}
   }
   \Else{
       Pick $\tilde{y}_t$ randomly from $
       \text{unif}[K]$ \label{line:explore}\\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t = y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} + x_t$ } \label{line:update2}
   }
}
\end{algorithm}
\textbf{Analysis}. Note that when the algorithm enters Line~\ref{line:update1} or Line~\ref{line:update2}, the binary classifier $w_t^{(\tilde{y}_t)}$ is making an error. Let the number of times the algorithm enters Line~\ref{line:update1} and Line~\ref{line:update2} be $M$ and $N$ respectively. By the error bound of binary perceptron $\mathcal{O}\left(\frac{1}{\gamma^2}\right)$, we have that $M+N=\mathcal{O}\left(\frac{K}{\gamma^2}\right)$. Then note that the number of times the algorithm makes a mistake (i.e. $\tilde{y}_t\neq y_t$) is upper bounded by $M$ plus how many times the algorithm explores in Line~\ref{line:explore}; and when the algorithm explores, with probability $\frac{1}{K}$ it enters Line~\ref{line:update2}. Therefore, the number of mistakes is bounded (in expectation) by $\E[M+KN]=\mathcal{O}\left(\frac{K^2}{\gamma^2}\right)$.

\subsection{One-versus-all Kernel Perceptron}
With the polynomial kernel defined in \cite{klivans2004learning}, we can transform the weaker assumption \eqref{eqn:weaker} in the original feature space to the stronger one \eqref{eqn:stronger} in a transformed feature space. Indeed, in the original feature space, class $i$ corresponds to an intersection of $K-1$ halfspaces: $ \{x: (w_i^*-w_j^*)^\top x \geq 0, \forall j\neq i\} $, which is the subject of \cite{klivans2004learning} (here, $w_i^{*\top}$ equals to $\e_i^\top W^*$ defined above). Directly using their kernel construction, we can have margin $ \left(\frac{\gamma}{d}\right)^{\Omega(K\log K\log (1/\gamma))} $ in the transformed feature space. Running kernelized version of Algorithm~\ref{alg:ova}, we can get $ K^2\left(\frac{d}{\gamma}\right)^{\mathcal{O}(K\log K\log (1/\gamma))} $ error bound directly.

\section{A Regret Lower Bound for A Certain Type of Algorithms}
\label{section: passive lower bound}
In this section, we try to construct an error lower bound for a certain type of algorithms. This type of algorithms does not make update when it makes a wrong prediction. For simplicity, we only consider binary classification.  More formally, the algorithms we consider satisfy the following assumption.

@@

\textbf{Discussion}. There is a construction such that $1/V$ can be of order $\Omega\left( \left(\frac{1}{\gamma}\right)^{(d-1)/2} \right)$.

%This lower bound does not rule out those algorithms that change its probability vector based on the $\textbf{count}$ of consecutive errors. This type of algorithm may still easy to be implemented (like QBC). Also it does not rule out the banditron algorithm.

%\section{Cone Algorithm}
%\begin{algorithm}[H]
%\caption{Banditron}
%\textbf{definition}: $K\triangleq$ number of classes, $\gamma\triangleq$  margin\\
%\textbf{Initialize}: $\calS_1=\cdots=\calS_K=\phi$ (empty set) \\
%\For{$t=1, \ldots, T$}{
%    Receive $x_t\in \mathbb{R}^d$. \\
%    Define the cone $\calC_i = \left\{x\in\mathbb{R}^d: x=\sum_{j=1}^{|\calS_i|} \alpha_jy_j, \text{\ where\ } y_j\in \calS_i, \alpha_j\geq 0 \right\}$ \\
%    (that is, $\calC_i$ is the conic hull of $\calS_i$) \\
%    Check whether $x_t$ belongs to, or has distance smaller than $\gamma$, to one of $\calC_1, \ldots, \calC_K$. \\
%    If so, classify $x_t$ to the corresponding class (say class $i$), and let $\calS_i\leftarrow \calS_i \cup \{x_t\}$. {\color{red} This prediction will be correct for sure by our margin assumption.}\\
%    If not, let $\tilde{y}_t\sim \text{unif}([K])$ and predict $\tilde{y}_t$. If $\tilde{y}_t=y_t$, then $\calS_{y_t}\leftarrow \calS_{y_t} \cup \{x_t\}$. 
%}
%\end{algorithm}

@@

@@

\newpage
\bibliography{adaptivity} 
\bibliographystyle{plain}
